The Semantics and Implementation of \CEU: a Synchronous Reactive Language based on Esterel

### 2015-09-19 paper.tex
Nova Secao 4.1 para descrever a analise temporal para acessso a memoria compratilhada.

@@

\item[Analysis]

This phase detects inconsistencies in \CEU programs, such as unbounded loops,
and also makes ``classical'' semantic analysis, such as building a symbol table 
for checking variable declarations.
%
This phase outputs code in $C$, given the way \CEU is tied to $C$ by design.
%
Some type checking is delayed to the last phase to take advantage of 
\code{gcc}'s error handling.
Therefore, we annotate the $C$ output file with \code{\#line} pragmas matching 
the original file in \CEU.

@@

\begin{comment}
\subsection{Temporal Analysis}

As introduced, the \emph{temporal analysis} phase detects inconsistencies in 
\CEU programs.
Here, we focus on the algorithm that detects non-deterministic access to 
variables, as presented in Section~\ref{sec.ceu.shared}.

@@

A node inherits the set $I$ from its direct parent and calculates $O$ according 
to its type:
%
\begin{itemize}
%
\item Nodes that represent expressions, assignments, $C$ calls, and 
declarations simply reproduce $O=I$, as they do not await;
%
\item An \code{await e} statement has $O=\{e\}$.
%
\item A \code{break} statement has $O=\{\}$ as it escapes the innermost 
\code{loop} and never terminate, i.e., never proceeds to the statement 
immediately following it (see also \code{loop} below);
%
\item A \emph{sequence node (;)} modifies each of its children to have 
$I_n=O_{n-1}$.
The first child inherits $I$ from the sequence parent, and the set $O$ for the 
sequence node is copied from its last child, i.e., $O=O_n$.
%
\item A \code{loop} node includes its body's $O$ on its own $I$ ($I=I \cup 
O_{body}$), as the loop is also reached from its own body.
The union of all \code{break} statements' $O$ forms the set $O$ for a 
\code{loop}.
%
\item An \code{if} node has $O=O_{true} \cup O_{false}$.
%
\item A parallel composition (\code{par/and} / \code{par/or}) may terminate 
from any of its branches, hence $O = O_1 \cup ... \cup O_n$.
\end{itemize}

With all sets calculated, any two nodes that perform side effects and are in 
parallel branches can have their $I$ sets compared for intersections.
If the intersection is not the empty set, they are marked as suspicious (see 
Section~\ref{sec.ceu.shared}).

Figure~\ref{lst.impl.ast} reproduces the second code of Figure~\ref{lst.det} 
and shows the corresponding $AST$ with the sets $I$ and $O$ for each node.
The event $.$ (dot) represents the ``boot'' reaction.
The assignments to \code{y} in parallel (lines 5,8 in the code) have an empty 
intersection of $I$ (lines 6,9 in the AST), hence, they do not conflict.
Note that although the accesses in lines 5, 11 in the code (lines 6,11 in the 
AST) do have an intersection, they are not in parallel and are also safe.

@@

begin{comment}
It is also responsible for setting the priorities for trails (see further) and 
determining the sizes of the queues that are used during runtime.

@@

From its starting node, the flow graph is traversed until reaching await 
nodes---every visited node is inserted into a new DFA state.
Then, every set of awaiting nodes for a given external event starts another DFA 
state.
end{comment}

\end{comment}

>>>>

% TODO: par/or vs abort, strong/weak abortion

@@

\item[Temporal Analysis]

This phase detects inconsistencies in \CEU programs, such as unbounded loops, 
suspicious accesses to shared memory, and also ``classical'' semantic analysis, 
such as building a symbol table for checking variable declarations.
%
This phase outputs code in $C$, given the way \CEU is tied to $C$ by design.
%
Some type checking is delayed to the last phase to take advantage of 
\code{gcc}'s error handling.
Therefore, we annotate the $C$ output file with \code{\#line} pragmas matching 
the original file in \CEU.

@@

\subsection{Temporal Analysis for Shared-Memory Concurrency}

The compile-time \emph{temporal analysis} phase detects inconsistencies in \CEU 
programs.
Here, we focus on the algorithm that detects suspicious access to shared 
variables, as discussed in Section~\ref{sec.ceu.det}.

@@

A node inherits the set $I$ from its direct parent and calculates $O$ according 
to its type:
%
\begin{itemize}
%
\item Nodes that represent expressions, assignments, $C$ calls, and 
declarations simply reproduce $O=I$, as they do not await;
%
\item An \code{await e} statement has $O=\{e\}$.
%
\item A \code{break} statement has $O=\{\}$ as it escapes the innermost 
\code{loop} and never terminates, i.e., never proceeds to the statement 
immediately following it (see also \code{loop} below);
%
\item A \emph{sequence node (;)} modifies each of its children to have 
$I_n=O_{n-1}$.
The first child inherits $I$ from the sequence parent, and the set $O$ for the 
sequence node is copied from its last child, i.e., $O=O_n$.
%
\item A \code{loop} node includes its body's $O$ on its own $I$ ($I=I \cup 
O_{body}$), as the loop is also reached from its own body.
The union of all \code{break} statements' $O$ forms the set $O$ for a 
\code{loop}.
%
\item An \code{if} node has $O=O_{true} \cup O_{false}$.
%
\item A parallel composition (\code{par/and} / \code{par/or}) may terminate 
from any of its branches, hence $O = O_1 \cup ... \cup O_n$.
\end{itemize}

With all sets calculated, any two nodes that perform side effects and are in 
parallel branches can have their $I$ sets compared for intersections.
If the intersection is not the empty set, they are marked as suspicious.

The example in the left of Figure~\ref{lst.impl.ast} has a corresponding $AST$, 
in the right of the figure, with the sets $I$ and $O$ for each node.
The event $.$ (dot) represents the ``boot'' reaction.
The assignments to \code{y} in parallel (lines 5,8 in the code) have an empty 
intersection of $I$ (lines 6,9 in the AST), hence, they do not conflict.
Note that although the accesses in lines 5,11 in the code (lines 6,11 in the 
AST) do have an intersection, they are not in parallel and are also safe.

@@

\begin{comment}
It is also responsible for setting the priorities for trails (see further) and 
determining the sizes of the queues that are used during runtime.

@@

From its starting node, the flow graph is traversed until reaching await 
nodes---every visited node is inserted into a new DFA state.
Then, every set of awaiting nodes for a given external event starts another DFA 
state.
\end{comment}
